|

### 2023-06-15 manuscript.qmd
author list updates

---
title: |
  Longitudinal Analysis Manuscript: Working Draft
subtitle: Longitudinal Analysis using data from the ABCD Study
short-title: Longitudinal Analysis
code-repo: "Access the code, data, and analysis at <http://swhawes.github.io/test/>"
number-sections: true
number-depth: 3
correspondence-prefix: "Correspondence concerning this article should be addressed to"
author:
  - name: Samuel Hawes
    email: shawes@fiu.edu
    title: Associate Research Professor
    affiliations:
      #- id: ds9
        name: Florida International University
        department: Center for Children & Families
        address: 11200 SW 8th St, AHC-IV
        city: Miami
        region: FL
        country: USA
        postal-code: 33199
    attributes:
      corresponding: true
  - name: Andrew K. Littlefield
    email: andrew.littlefield@ttu.edu
    affiliations:
      #- id: ds9
        name: Texas Tech University
        department: Psychological Sciences
        city: Lubbock
        region: TX
        country: USA
  - name: Daniel A. Lopez
    email: Daniel_Lopez2@URMC.Rochester.edu
    affiliations:
      #- id: terok
        name: University of Rochester Medical Center
        department: Department of Public Health
        city: Rochester
        region: NY
        country: USA
      #- ref: ds9
  - name: Kenneth J. Sher
    email: SherK@missouri.edu
    affiliations:
      #- id: ds9
        name: University of Missouri
        department: Psychological Sciences
        city: Columbia
        region: MO
        country: USA
  # - name: Erin L. Thompson
  #   email: erthomps@fiu.edu
  #   affiliations:
  #     #- id: terok
  #       name: Florida International University
  #       department: Center for Children and Families
  #       city: Miami
  #       region: FL
  #       country: USA
  # - name: Raul Gonzalez
  #   email: gonzara@fiu.edu
  #   affiliations:
  #     #- id: ds9
  #       name: Florida International University
  #       department: Center for Children and Families
  #       city: Miami
  #       region: FL
  #       country: USA
  - name: Additional Co-authors
    email: xxxxx@fiu.edu
    affiliations:
      #- id: terok
        name: xxxxxx University
        department: xxxxxx Dept
        city: xxxx
        region: xxxxx
        country: xxxxx
  - name: Wesley K. Thompson
    email: wes.stat@gmail.com
    affiliations:
      #- id: terok
        name: Laureate Institute for Brain Research
        department: Center for Population Neuroscience and Genetics
        city: Tulsa
        region: OK
        country: USA

@@

\label{sec:headings} [The Adolescent Brain Cognitive Development (ABCD) Study®](https://abcdstudy.org/ "Visit the ABCD Study website"){#abcdwebsite} is the largestz long-term investigation of neurodevelopment and child health in the United States. Conceived and initiated by the National Institutes of Health (NIH), this landmark prospective longitudinal study aims to transform our understanding of the genetic and environmental influences on brain development and their roles in behavioral and health outcomes in adolescents [@volkow2018]. At its heart, the study is designed to chart the course of human development across multiple, interacting domains from late childhood to early adulthood and to identify factors that lead to both positive and negative developmental outcomes. Central to achieving these goals is the ABCD Study's® commitment to an open science framework designed to facilitate access to and sharing of scientific knowledge by espousing practices that increase openness, integrity, and reproducibility of scientific research (e.g., public data releases). In this sense, the ABCD Study® is a collaboration with the larger research community, with the rich longitudinal nature of the ABCD Study dataset allowing researchers to perform a variety of analyses of both methodological and substantive interest. Together, this presents a unique opportunity to significantly advance our understanding of how a multitude of biopsychosocial processes emerge and unfold across critical periods of development. In the following text, we describe methods for longitudinal analyses of ABCD Study data that can address its fundamental scientific aims, as well as challenges inherent in a large population-based long-term study of adolescents. We begin with a brief orientation to the ABCD Study and its accompanying [publicly available dataset](https://nda.nih.gov/abcd/ "Visit the ABCD Study data repository"){#ndardata}.

>>>>

---
title: |
  Longitudinal Analysis Manuscript: Working Draft
subtitle: Longitudinal Analysis using data from the ABCD Study
short-title: Longitudinal Analysis
code-repo: "Access the code, data, and analysis at <http://swhawes.github.io/test/>"
number-sections: true
number-depth: 3
correspondence-prefix: "Correspondence concerning this article should be addressed to"
author:
  - name: Samuel Hawes
    email: shawes@fiu.edu
    title: Associate Research Professor
    affiliations:
      #- id: ds9
        name: Florida International University
        department: Center for Children & Families
        address: 11200 SW 8th St, AHC-IV
        city: Miami
        region: FL
        country: USA
        postal-code: 33199
    attributes:
      corresponding: true
  - name: Andrew K. Littlefield
    email: andrew.littlefield@ttu.edu
    affiliations:
      #- id: ds9
        name: Texas Tech University
        department: Psychological Sciences
        city: Lubbock
        region: TX
        country: USA
  - name: Daniel A. Lopez
    email: Daniel_Lopez2@URMC.Rochester.edu
    affiliations:
      #- id: terok
        name: University of Rochester Medical Center
        department: Department of Public Health
        city: Rochester
        region: NY
        country: USA
      #- ref: ds9
  - name: Kenneth J. Sher
    email: SherK@missouri.edu
    affiliations:
      #- id: ds9
        name: University of Missouri
        department: Psychological Sciences
        city: Columbia
        region: MO
        country: USA
  - name: Erin L. Thompson
    email: erthomps@fiu.edu
    affiliations:
      #- id: terok
        name: Florida International University
        department: Center for Children and Families
        city: Miami
        region: FL
        country: USA
  - name: Raul Gonzalez
    email: gonzara@fiu.edu
    affiliations:
      #- id: ds9
        name: Florida International University
        department: Center for Children and Families
        city: Miami
        region: FL
        country: USA
  - name: Additional Co-authors
    email: xxxxx@fiu.edu
    affiliations:
      #- id: terok
        name: xxxxxx University
        department: xxxxxx Dept
        city: xxxx
        region: xxxxx
        country: xxxxx
  - name: Wesley K. Thompson
    email: wes.stat@gmail.com
    affiliations:
      #- id: terok
        name: Laureate Institute for Brain Research
        department: Center for Population Neuroscience and Genetics
        city: Tulsa
        region: OK
        country: USA

@@

\label{sec:headings} [The Adolescent Brain Cognitive Development (ABCD) Study®](https://abcdstudy.org/ "Visit the ABCD Study website"){#abcdwebsite} is the largest long-term investigation of neurodevelopment and child health in the United States. Conceived and initiated by the National Institutes of Health (NIH), this landmark prospective longitudinal study aims to transform our understanding of the genetic and environmental influences on brain development and their roles in behavioral and health outcomes in adolescents [@volkow2018]. At its heart, the study is designed to chart the course of human development across multiple, interacting domains from late childhood to early adulthood and to identify factors that lead to both positive and negative developmental outcomes. Central to achieving these goals is the ABCD Study's® commitment to an open science framework designed to facilitate access to and sharing of scientific knowledge by espousing practices that increase openness, integrity, and reproducibility of scientific research (e.g., public data releases). In this sense, the ABCD Study® is a collaboration with the larger research community, with the rich longitudinal nature of the ABCD Study dataset allowing researchers to perform a variety of analyses of both methodological and substantive interest. Together, this presents a unique opportunity to significantly advance our understanding of how a multitude of biopsychosocial processes emerge and unfold across critical periods of development. In the following text, we describe methods for longitudinal analyses of ABCD Study data that can address its fundamental scientific aims, as well as challenges inherent in a large population-based long-term study of adolescents. We begin with a brief orientation to the ABCD Study and its accompanying [publicly available dataset](https://nda.nih.gov/abcd/ "Visit the ABCD Study data repository"){#ndardata}.
